---
title: "Take-home_Ex02"
format:
  html:
    toc: true
    toc-location: left
    code-fold: true
    code-summary: "Show the code"
execute:
  warning: false
  message: false
---
## Overview

TenantThread is an AI-agent system that manages a property platform's public and internal communications. Seven autonomous agents — covering legal, platform-trust, PR, social-media and compliance roles — coordinate across six channels, with a dedicated "Judge" agent reviewing official posts for compliance.

On 5 June 2046, material about the confidential CivicLoom acquisition (codenamed *Project HarborCrest*) — under embargo until 18:00 that day — surfaced publicly before the embargo lifted. This report investigates a single forensic question: **was the leak a deliberate, coordinated act, or did the system simply break down under pressure?**

To answer it, the analysis works through three questions — reconstructing what happened (Q1), contrasting normal behaviour against the breach (Q2), and asking whether warning signs appeared earlier and why no one acted (Q3) — using interactive network and text-analysis visualisations throughout.

## Data and methodology

::: {.callout-note}
## How to read these visualisations

Two ideas recur throughout the report:

- **The oversight ladder.** Every public channel is ranked by how much compliance scrutiny it carries, encoded by colour from safe to risky: **Internal** (grey, never published) → **Official** (green, reviewed by the Judge) → **Personal** (orange, low oversight) → **Anonymous** (red, no oversight). Reading a figure top-to-bottom or grey-to-red is reading *down the oversight ladder*.
- **Phase.** Rounds are split into **Pre-crisis** (17 May – 4 June, the system's normal state) and the **Crisis day** (5 June), so every view can be compared against the baseline.

The forensic question — deliberate leak versus breakdown under pressure — turns on a simple test. A **deliberate** act would concentrate cleanly on the unmonitored channels and show foresight in the agents' private reasoning. A **breakdown** would be diffuse and unplanned, with no advance signal. Each figure is read against that expectation.
:::

The dataset is a nested JSON file (`MC1_final_00.json`) of 23 communication rounds: 13 daily rounds across the pre-crisis period (17 May – 4 June) and 10 hourly rounds on the crisis day (5 June, 09:00–18:00), totalling 912 messages between the seven agents.

Each message carries a sender, channel, recipients, timestamp and content, plus the agent's internal state (its private deliberation, declared action and self-reported urgency). During preparation the nested structure was flattened into tidy tables — messages, participants, edges and a deliberation corpus — and enriched with derived fields: a four-level **oversight** scale (from internal, through Judge-reviewed official posts, down to no-oversight anonymous posts) and a **phase** marker (pre-crisis vs crisis day).

Two analytical lenses are combined. **Network analysis** (tidygraph / ggraph) maps who communicated with whom, how the structure shifted, and who became central. **Text analysis** (tidytext) examines what the agents said and privately deliberated. All key figures are interactive (ggiraph, visNetwork) so findings can be explored directly.
```{r}
pacman::p_load(
  jsonlite,      # parse the nested JSON
  tidyverse,     # dplyr / tidyr / purrr / ggplot2 / stringr
  lubridate,     # date-time handling
  ggtext,        # rich text in titles (markdown-style)
  knitr,         # kable() tables
  patchwork,     # combine ggplots
  tidygraph,     # graph object + dplyr verbs on graphs
  ggraph,        # network plotting
  igraph,        # network algorithms (centrality, etc.)
  ggiraph,       # interactive hover plots
  tidytext,      # text mining (for Q2 / Q3)
  widyr,         # pairwise word correlation (Q3)
  ggwordcloud,   # word clouds (Q2)
  DT,            # interactive tables
  visNetwork
)
```


```{r}
# ---- Load data ----
mc1_raw <- fromJSON("data/MC1_final_00.json", simplifyVector = FALSE)
rounds  <- mc1_raw$rounds
length(rounds)   # expect 23

# ---- Central style definitions (single source of truth) ----
STYLES <- list(
  font_family = "sans",
  # oversight ramp: grey = safe/internal  ->  red = no oversight
  oversight_colors = c(
    "Internal"                  = "#B4B2A9",
    "Official (Judge-reviewed)" = "#1D9E75",
    "Personal (low oversight)"  = "#EF9F27",
    "Anonymous (no oversight)"  = "#E24B4A"
  ),
  # seniority ramp (echoes the example's supertype encoding)
  seniority_colors = c(
    "Senior"     = "#378ADD",
    "Junior"     = "#BA7517",
    "Compliance" = "#7F77DD"
  ),
  node_size         = 8,
  node_border_color = "grey30",
  edge_color        = "grey75",
  title_color       = "grey15",
  tooltip_css = paste0(
    "background-color:#1a1a1a;color:#fff;",
    "font-family:sans-serif;font-size:10pt;",
    "padding:6px;border-radius:4px;"
  ),
  svg_width  = 7,
  svg_height = 7 * 0.618
)
```

```{r}
`%||%` <- function(x, y) if (is.null(x)) y else x
```

```{r}
# ---- Flatten rounds -> communications into one row per message ----
messages <- map_dfr(rounds, function(rd) {
  hour <- rd$hour
  map_dfr(rd$communications, function(m) {
    ist <- m$internal_state %||% list()
    tibble(
      message_id    = m$message_id    %||% NA_character_,
      hour          = hour            %||% NA_character_,
      timestamp     = m$timestamp     %||% NA_character_,
      agent_id      = m$agent_id      %||% NA_character_,
      agent_role    = m$agent_role    %||% NA_character_,
      agent_label   = m$agent_label   %||% NA_character_,
      channel       = m$channel       %||% NA_character_,
      message_type  = m$message_type  %||% NA_character_,
      responding_to = m$responding_to %||% NA_character_,
      recipients    = paste(unlist(m$recipients), collapse = ";"),
      content       = m$content       %||% NA_character_,
      reacting      = ist$reacting      %||% NA_character_,
      rationalizing = ist$rationalizing %||% NA_character_,
      deliberating  = ist$deliberating  %||% NA_character_
    )
  })
})
```

```{r}
# ---- Role -> seniority lookup (from the data description) ----
seniority_lookup <- c(
  legal = "Senior", platform_trust = "Senior", pr = "Senior",
  social_media = "Senior", intern = "Junior", pr_intern = "Junior",
  judge = "Compliance"
)
```

```{r}
# ---- Add classification fields (our analog of supertype/subtype) ----
messages <- messages %>%
  mutate(
    timestamp = lubridate::ymd_hms(timestamp),
    date      = lubridate::as_date(timestamp),
    phase     = factor(if_else(date < lubridate::as_date("2046-06-05"),
                               "Pre-crisis", "Crisis day"),
                       levels = c("Pre-crisis", "Crisis day")),
    oversight = factor(case_when(
      channel == "official_post"  ~ "Official (Judge-reviewed)",
      channel == "personal_post"  ~ "Personal (low oversight)",
      channel == "anonymous_post" ~ "Anonymous (no oversight)",
      TRUE                        ~ "Internal"
    ), levels = names(STYLES$oversight_colors)),
    is_public = channel %in% c("official_post", "personal_post", "anonymous_post"),
    seniority = unname(seniority_lookup[agent_role])
  )
```

```{r}
# ---- Participants: declared action + sentiment + action class per round ----
participants <- map_dfr(rounds, function(rd) {
  hour <- rd$hour
  map_dfr(rd$participants, function(p) {
    meta <- p$agent_round_metadata %||% list()
    tibble(
      hour                  = hour %||% NA_character_,
      agent_id              = p$agent_id    %||% NA_character_,
      agent_role            = p$agent_role  %||% NA_character_,
      agent_label           = p$agent_label %||% NA_character_,
      declared_action       = p$declared_action %||% NA_character_,
      sentiment_at_turn     = meta$sentiment_at_turn %||% NA_character_,
      action_classification = meta$action_classification %||% NA_character_
    )
  })
}) %>%
  mutate(timestamp = lubridate::ymd_hms(hour), date = lubridate::as_date(timestamp))
```

```{r}
# ---- Environment context per round (for the Q1 timeline narrative) ----
env <- map_dfr(rounds, function(rd) {
  ec <- rd$environment_context %||% list()
  ms <- ec$market_snapshot      %||% list()
  tibble(
    hour            = rd$hour            %||% NA_character_,
    event_headline  = ec$event_headline  %||% NA_character_,
    event_narrative = ec$event_narrative %||% NA_character_,
    social_state    = ec$social_state    %||% NA_character_,
    stock_price     = ms$stock_price     %||% NA_character_,
    sentiment       = ms$sentiment       %||% NA_character_
  )
}) %>%
  mutate(timestamp = lubridate::ymd_hms(hour))

nrow(participants)   # expect 100
nrow(env)            # expect 23
```

```{r}
# ---- Node metadata table (name / label / alias / seniority) ----
agent_nodes <- messages %>%
  distinct(agent_role, agent_label) %>%
  mutate(
    seniority = unname(seniority_lookup[agent_role]),
    alias = recode(agent_role,
      legal = "LEG", platform_trust = "TRU", pr = "PR",
      social_media = "SOC", intern = "INT", pr_intern = "PRI", judge = "JDG")
  ) %>%
  rename(name = agent_role, label = agent_label) %>%
  arrange(name)
```

```{r}
# ---- Build directed edges (agent -> recipient) ----
fix_role  <- function(x) recode(x, "social_manager" = "social_media")
all_roles <- agent_nodes$name

edges_long <- messages %>%
  filter(recipients != "") %>%                 # drop public posts / external DMs
  separate_rows(recipients, sep = ";") %>%
  rename(recipient = recipients) %>%
  mutate(recipient = fix_role(recipient))
```

```{r}
# broadcasts ("ALL") -> expand to every other role
broadcast <- edges_long %>%
  filter(recipient == "ALL") %>%
  mutate(to = list(all_roles)) %>%
  unnest(to) %>%
  filter(to != agent_role)
```

```{r}
# targeted messages -> the named recipient
targeted <- edges_long %>%
  filter(recipient != "ALL") %>%
  rename(to = recipient) %>%
  filter(to != agent_role)

edges_all <- bind_rows(broadcast, targeted) %>%
  transmute(from = agent_role, to, channel, message_type, phase, date)

edges_agg <- edges_all %>%
  group_by(from, to) %>%
  summarise(weight = n(), .groups = "drop")
```

```{r}
# ---- The supernetwork (like the example's tbl_graph) ----
supernetwork <- tbl_graph(nodes = agent_nodes, edges = edges_agg, directed = TRUE)
supernetwork                                   # expect 7 nodes, a directed multigraph
```

```{r}
# ---- Text tables for Q2 / Q3 (tidytext) ----
text_df <- messages %>%
  select(message_id, agent_label, agent_role, phase, date,
         channel, oversight, content) %>%
  filter(!is.na(content), content != "")

tidy_words <- text_df %>%
  unnest_tokens(word, content) %>%
  filter(!word %in% stop_words$word, str_detect(word, "[a-z]"))

delib_df <- messages %>%
  filter(!is.na(deliberating)) %>%
  select(message_id, agent_label, phase, date, deliberating)
```

```{r}
# ---- Save everything as rds (compute once, reuse later) ----
if (!dir.exists("data/rds")) dir.create("data/rds", recursive = TRUE)
write_rds(messages,     "data/rds/messages.rds")
write_rds(participants, "data/rds/participants.rds")
write_rds(env,          "data/rds/env.rds")
write_rds(agent_nodes,  "data/rds/agent_nodes.rds")
write_rds(edges_all,    "data/rds/edges_all.rds")
write_rds(edges_agg,    "data/rds/edges_agg.rds")
write_rds(supernetwork, "data/rds/supernetwork.rds")
write_rds(text_df,      "data/rds/text_df.rds")
write_rds(tidy_words,   "data/rds/tidy_words.rds")
write_rds(delib_df,     "data/rds/delib_df.rds")
```

```{r}
# ---- Sanity check ----
cat("rounds:", length(rounds), "| messages:", nrow(messages),
    "| public:", sum(messages$is_public), "| participants:", nrow(participants),
    "| env:", nrow(env), "| nodes:", nrow(agent_nodes), "\n")
```


## The communication network at a glance

Before the analysis, this interactive view lets you explore TenantThread's seven agents and how they communicate. Drag a node, zoom, or select an agent from the dropdown to highlight its connections. Node colour shows seniority; node size reflects how many messages each agent sent.

```{r}
# ---- (Showcase) Fully interactive communication network (visNetwork) ----
agent_nodes <- read_rds("data/rds/agent_nodes.rds")
edges_all   <- read_rds("data/rds/edges_all.rds")

# messages sent per agent -> drives node size
agent_msgs <- read_rds("data/rds/messages.rds") %>%
  count(agent_role, name = "msgs") %>%
  rename(name = agent_role)
```

```{r}
# nodes: visNetwork needs id (matching edges) + label/group/value/title
vis_nodes <- agent_nodes %>%
  left_join(agent_msgs, by = "name") %>%
  mutate(
    msgs  = tidyr::replace_na(msgs, 0L),
    id    = name,
    group = seniority,
    value = msgs,
    title = paste0("<b>", label, "</b><br>Seniority: ", seniority,
                   "<br>Messages sent: ", msgs)
  ) %>%
  select(id, label = alias, group, value, title)
```

```{r}
# edges: agent-to-agent ties only; weight = number of messages
vis_edges <- edges_all %>%
  filter(from %in% vis_nodes$id, to %in% vis_nodes$id) %>%
  count(from, to, name = "weight") %>%
  mutate(
    value  = weight,
    title  = paste0(from, " &rarr; ", to, ": ", weight, " messages"),
    arrows = "to"
  )

visNetwork(vis_nodes, vis_edges,
           main    = "Explore the agent communication network",
           submain = "Drag nodes, zoom, or pick an agent to highlight its ties") %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visGroups(groupname = "Senior",
            color = unname(STYLES$seniority_colors["Senior"])) %>%
  visGroups(groupname = "Junior",
            color = unname(STYLES$seniority_colors["Junior"])) %>%
  visGroups(groupname = "Compliance",
            color = unname(STYLES$seniority_colors["Compliance"])) %>%
  visEdges(color  = list(color = "#cccccc", highlight = "#E24B4A"),
           smooth = list(enabled = TRUE, type = "curvedCW")) %>%
  visOptions(highlightNearest = list(enabled = TRUE, degree = 1, hover = TRUE),
             nodesIdSelection = TRUE) %>%
  visLegend(main = "Seniority") %>%
  visLayout(randomSeed = 1234)
```
::: {.callout-tip}
## Insights

Even before any time analysis, the shape is suggestive. The four **Senior** agents — legal, platform-trust, PR and social-media — are the large, densely connected nodes that carry most of the traffic, while the **Compliance** agent (the Judge) is comparatively small and loosely tied in. Whoever drives the system's communication, it is not the compliance function.
:::


## Q1 — Reconstructing the breach

This first question rebuilds the event itself: what the system normally looked like, what changed on the crisis day, and which agents drove that change. The three figures move from *what was posted* (public-post volume) to *who coordinated it* (private network) to *when it happened* (crisis-day timeline).

```{r}
# ---- Q1 setup: load prepared data + a shared ggplot theme ----
messages <- read_rds("data/rds/messages.rds")
env      <- read_rds("data/rds/env.rds")

# Reusable theme that pulls from STYLES (single source of truth)
theme_mc1 <- function() {
  theme_minimal(base_family = STYLES$font_family) +
    theme(
      plot.title       = element_text(color = STYLES$title_color, face = "bold"),
      plot.subtitle    = element_text(color = "grey40", size = 10),
      legend.position  = "bottom",
      panel.grid.minor = element_blank(),
      axis.text.x      = element_text(angle = 45, hjust = 1)
    )
}
```

### Figure 1 — Public posts per round, by oversight

We start with the raw output of the system: how many public posts each round produced, stacked by where they sat on the oversight ladder.
```{r}
# ---- Q1, Figure 1 (revised): public-post volume per ROUND, faceted by phase ----
posts_by_round <- messages %>%
  filter(is_public) %>%
  count(hour, phase, oversight, name = "n") %>%
  mutate(
    round_time = lubridate::ymd_hms(hour),
    # short, phase-aware label: date before the crisis, clock time on the crisis day
    round_lbl  = if_else(phase == "Crisis day",
                         format(round_time, "%H:%M"),
                         format(round_time, "%b %d")),
    round_lbl  = fct_reorder(round_lbl, round_time),
    tooltip    = paste0(oversight, "<br/>",
                        format(round_time, "%b %d, %H:%M"), "<br/>",
                        n, " public post(s)")
  )

p1 <- ggplot(posts_by_round,
             aes(x = round_lbl, y = n, fill = oversight)) +
  geom_col_interactive(aes(tooltip = tooltip, data_id = round_lbl),
                       width = 0.85) +
  scale_fill_manual(values = STYLES$oversight_colors, drop = FALSE) +
  facet_grid(~ phase, scales = "free_x", space = "free_x") +
  labs(
    title    = "Public posts per round, by channel oversight",
    subtitle = "Anonymous (no-oversight) posts surface only on the crisis day, June 5",
    x = NULL, y = "Public posts", fill = "Channel oversight"
  ) +
  theme_mc1() +
  theme(panel.spacing = unit(1, "lines"))

girafe(ggobj   = p1,
       width_svg  = STYLES$svg_width,
       height_svg = STYLES$svg_height,
       options = list(opts_tooltip(css = STYLES$tooltip_css)))
```
::: {.callout-tip}
## Insights

Across all thirteen pre-crisis rounds, public posting is sparse and runs only through gated or low-oversight channels — the **Anonymous** (red, no-oversight) channel is never touched. It surfaces for the first time on **June 5**, where the bars are also taller and lean increasingly red toward the embargo. The no-oversight channel is, on this evidence, a crisis-day artefact.
:::

### Figure 2 — Private coordination network

Public posts show *what surfaced*; the private channels show *who arranged it*. This network keeps only one-on-one chats and side-huddles — the conversations that never reach a public feed.
```{r}
# ---- Q1, Figure 2: private-coordination network (interactive) ----
# 1) node attributes: activity (size) + who pushed the breach (highlight)
agent_stats <- messages %>%
  count(agent_role, name = "msgs_sent") %>%
  rename(name = agent_role)

leak_actors <- messages %>%                       # used low/no-oversight public channels on June 5
  filter(phase == "Crisis day",
         channel %in% c("personal_post", "anonymous_post")) %>%
  distinct(agent_role) %>%
  pull(agent_role)

node_tbl <- read_rds("data/rds/agent_nodes.rds") %>%
  left_join(agent_stats, by = "name") %>%
  mutate(
    msgs_sent = tidyr::replace_na(msgs_sent, 0L),
    emphasis  = name %in% leak_actors | name == "judge",
    tooltip   = paste0(label, "<br/>Seniority: ", seniority,
                       "<br/>Messages sent: ", msgs_sent)
  )

# 2) keep only PRIVATE/targeted edges -> reveals coordination structure
edges_private <- read_rds("data/rds/edges_all.rds") %>%
  filter(channel %in% c("one_on_one_chat", "side_huddle")) %>%
  count(from, to, name = "weight")

g_priv <- tbl_graph(nodes = node_tbl, edges = edges_private, directed = TRUE)
```

```{r}
set.seed(1234)
p2 <- ggraph(g_priv, layout = "fr") +
  geom_edge_fan(aes(width = weight),
                color = STYLES$edge_color, alpha = 0.55,
                arrow = arrow(length = unit(2, "mm"), type = "closed"),
                end_cap = circle(4, "mm"), show.legend = FALSE) +
  scale_edge_width(range = c(0.3, 2.2)) +
  geom_point_interactive(
    aes(x = x, y = y, size = msgs_sent, color = seniority,
        tooltip = tooltip, data_id = name)) +
  geom_point(data = ~ dplyr::filter(.x, emphasis),          # black ring on key actors
             aes(x = x, y = y, size = msgs_sent),
             shape = 21, fill = NA, color = "black", stroke = 1.3,
             show.legend = FALSE) +
  geom_node_text(aes(label = alias),
                 color = "white", fontface = "bold", size = 3) +
  scale_color_manual(values = STYLES$seniority_colors) +
  scale_size(range = c(5, 13), guide = "none") +
  labs(title    = "Private coordination network (one-on-one & side huddles)",
       subtitle = "Black ring = posted via a low/no-oversight channel on June 5, or is the Judge",
       color    = "Seniority") +
  theme_graph(base_family = STYLES$font_family) +
  theme(legend.position = "bottom")

girafe(ggobj = p2,
       width_svg  = STYLES$svg_width,
       height_svg = STYLES$svg_height,
       options = list(opts_tooltip(css = STYLES$tooltip_css)))
```
::: {.callout-tip}
## Insights

Private coordination clusters tightly around the Senior agents (legal, platform-trust, social-media, PR). The agents who later published through personal or anonymous channels on June 5, together with the Judge, are ringed in black: the would-be leakers sit at the centre of the private traffic, while the **Judge sits on the periphery with few private ties** — structurally outside the loop it is meant to police.
:::

### Figure 3 — Crisis-day timeline

Finally we place every crisis-day public post on a clock, against the 6 PM embargo, to see exactly how the timing worked.
```{r}
# ---- Q1, Figure 3: crisis-day timeline - how the embargo was bypassed ----
crisis_posts <- messages %>%
  filter(phase == "Crisis day", is_public) %>%
  mutate(
    oversight = forcats::fct_drop(oversight),          # drop the empty "Internal" level
    tooltip   = paste0(format(timestamp, "%H:%M"), " - ", agent_label, "<br/>",
                       channel, "<br/>", str_trunc(content, 110))
  )

embargo <- lubridate::ymd_hms("2046-06-05 18:00:00")   # 6 PM embargo lift

p3 <- ggplot(crisis_posts, aes(x = timestamp, y = oversight, color = oversight)) +
  geom_vline(xintercept = embargo, linetype = "dashed", color = "grey40") +
  geom_point_interactive(aes(tooltip = tooltip, data_id = message_id),
                         size = 3.5, alpha = 0.9,
                         position = position_jitter(height = 0.12, width = 0)) +
  scale_color_manual(values = STYLES$oversight_colors, drop = FALSE, guide = "none") +
  scale_x_datetime(date_breaks = "1 hour", date_labels = "%H:%M") +
  labs(title    = "Crisis-day timeline: how the embargo was bypassed",
       subtitle = "Dashed line = 6 PM embargo lift. Low/no-oversight posts fire before it; the Judge never gates them.",
       x = "June 5", y = NULL) +
  theme_mc1()

girafe(ggobj = p3,
       width_svg  = STYLES$svg_width,
       height_svg = STYLES$svg_height,
       options = list(opts_tooltip(css = STYLES$tooltip_css)))
```
::: {.callout-tip}
## Insights

The low- and no-oversight posts (orange and red) fire steadily through the afternoon and bunch up just **before the dashed 18:00 embargo line**, while Judge-reviewed official posts (green) stay sparse throughout. The embargo was not broken by accident of timing — it was bypassed by a deliberate combination of *when* (before 6 PM) and *through what* (channels the Judge never gates).
:::

Q1 — Sequence of events and the embargo bypass
For the two weeks before the breach, TenantThread's agents published almost only through channels with oversight: official posts, which "The Judge" reviews, plus a few personal posts. The no-oversight anonymous channel was never used (Figure 1) — this was the system's normal, gated state.
June 5 broke the pattern. Public posting surged and the anonymous channel was activated for the first time; personal and anonymous posts rose through the afternoon and peaked at 5–6 PM (Figures 1 and 3). Critically, these low- and no-oversight posts went out before the 6 PM embargo lift, while Judge-reviewed official posts stayed sparse (Figure 3).
The private-coordination network shows who drove this: the legal, platform-trust, social-media and PR agents were the central, active nodes in one-on-one and side-huddle exchanges, whereas the Judge sat on the periphery (Figure 2). On the day, the Judge sent only six messages, all internal — it gated no personal or anonymous post.
The decisive system element is a coverage gap: the Judge enforces the embargo only on official posts, so embargoed content routed through personal and anonymous channels met no compliance check. The agents' choice to publish through exactly those unmonitored channels — including an anonymous channel switched on only that day — indicates at least partial intent. Whether this was a coordinated leak or a breakdown under pressure is tested in Q2 and Q3.

## Q2 — Normal behaviour versus the breach

Having reconstructed the event, Q2 contrasts the two weeks of normal operation against the crisis day along three axes — *which channels carried traffic*, *how the network re-centred*, and *how the language changed* — to test whether June 5 was simply busier, or structurally different.

```{r}
# ---- Q2, Figure 1: channel activity heatmap over time ----
messages <- read_rds("data/rds/messages.rds")

# fixed channel order: internal channels on top, public/risky channels at the bottom-up
channel_levels <- c("comms_huddle", "side_huddle", "one_on_one_chat",
                    "official_post", "personal_post", "anonymous_post")

heat <- messages %>%
  count(hour, phase, channel, name = "n") %>%
  # fill in every channel x round cell with 0 so empty cells show as blank
  complete(nesting(hour, phase), channel = channel_levels, fill = list(n = 0)) %>%
  mutate(
    round_time = lubridate::ymd_hms(hour),
    round_lbl  = if_else(phase == "Crisis day",
                         format(round_time, "%H:%M"),
                         format(round_time, "%b %d")),
    round_lbl  = fct_reorder(round_lbl, round_time),
    channel    = factor(channel, levels = channel_levels),
    tooltip    = paste0(channel, "<br/>", format(round_time, "%b %d %H:%M"),
                        "<br/>", n, " message(s)")
  )
```
### Figure 1 — Channel activity over time

To compare the baseline against the breach, we first lay out every channel against every round as a heatmap, ordered down the oversight ladder.
```{r}
p_heat <- ggplot(heat, aes(x = round_lbl, y = channel, fill = n)) +
  geom_tile_interactive(aes(tooltip = tooltip, data_id = paste(round_lbl, channel)),
                        color = "white", linewidth = 0.3) +
  scale_fill_gradient(low = "grey92", high = "#1D4E89",
                      trans = "sqrt", name = "Messages") +
  facet_grid(~ phase, scales = "free_x", space = "free_x") +
  labs(title    = "Channel activity over time",
       subtitle = "The anonymous (no-oversight) channel is used for the first time on the crisis day",
       x = NULL, y = NULL) +
  theme_mc1() +
  theme(panel.grid = element_blank())

girafe(ggobj = p_heat,
       width_svg  = STYLES$svg_width,
       height_svg = STYLES$svg_height,
       options = list(opts_tooltip(css = STYLES$tooltip_css)))
```
::: {.callout-tip}
## Insights

Internal channels carry steady, unremarkable traffic for the whole fortnight. The **anonymous row is completely empty across all thirteen pre-crisis rounds** and lights up only on June 5 — the single clearest signal that the crisis day is not "more of the same" but a categorical break from normal operation.
:::

### Figure 2 — Communication network: before versus during

Placing the two phases side by side — on an identical layout seed, so positions are comparable — shows how the structure itself re-wired.
```{r}
# ---- Q2, Figure 2: Pre-crisis vs Crisis day network (side-by-side) ----

# Load required data
edges_all   <- read_rds("data/rds/edges_all.rds")
agent_nodes <- read_rds("data/rds/agent_nodes.rds")

# Aggregate edge weight by phase (all channels combined)
edges_phase <- edges_all %>%
  count(from, to, phase, name = "weight")

# Per-phase message count per agent (drives node size)
node_phase_stats <- messages %>%
  count(agent_role, phase, name = "msgs_sent") %>%
  rename(name = agent_role)

```

```{r}
# Helper: build tbl_graph for one phase
make_phase_graph <- function(ph) {
  e <- edges_phase %>%
    filter(phase == ph) %>%
    select(from, to, weight)
  n <- agent_nodes %>%
    left_join(filter(node_phase_stats, phase == ph), by = "name") %>%
    mutate(
      msgs_sent = replace_na(msgs_sent, 0L),
      tooltip   = paste0("<b>", label, "</b><br/>",
                         "Phase: ", ph, "<br/>",
                         "Messages sent: ", msgs_sent, "<br/>",
                         "Seniority: ", seniority)
    )
  tbl_graph(nodes = n, edges = e, directed = TRUE)
}

g_pre    <- make_phase_graph("Pre-crisis")
g_crisis <- make_phase_graph("Crisis day")
```

```{r}
# Helper: plot one phase network (same seed = comparable layout)
plot_net <- function(g, title_str, show_legend = FALSE) {
  set.seed(42)
  ggraph(g, layout = "fr") +
    geom_edge_fan(
      aes(width = weight, alpha = weight),
      color = STYLES$edge_color,
      arrow = arrow(length = unit(2, "mm"), type = "closed"),
      end_cap = circle(4.5, "mm"), show.legend = FALSE
    ) +
    scale_edge_width(range = c(0.4, 2.5)) +
    scale_edge_alpha(range = c(0.25, 0.75), guide = "none") +
    geom_point_interactive(
      aes(x = x, y = y,
          size    = msgs_sent,
          color   = seniority,
          tooltip = tooltip,
          data_id = name),
      alpha = 0.95
    ) +
    geom_node_text(
      aes(label = alias),
      color = "white", fontface = "bold", size = 3.2
    ) +
    scale_color_manual(values = STYLES$seniority_colors) +
    scale_size(range = c(6, 14), guide = "none") +
    labs(title = title_str, color = "Seniority") +
    theme_graph(base_family = STYLES$font_family) +
    theme(
      legend.position = if (show_legend) "bottom" else "none",
      plot.title = element_text(size = 11, face = "bold",
                                color = STYLES$title_color)
    )
}

p_net_pre    <- plot_net(g_pre,    "Pre-crisis  (May 17 – Jun 4)")
p_net_crisis <- plot_net(g_crisis, "Crisis day  (Jun 5)",
                         show_legend = TRUE)

p_net2 <- p_net_pre + p_net_crisis +
  plot_annotation(
    title    = "Communication network: Pre-crisis vs Crisis day",
    subtitle = "Node size ∝ messages sent in that phase · edges = all channels · hover for detail",
    theme    = theme(
      plot.title    = element_text(face = "bold", size = 13,
                                   color = STYLES$title_color),
      plot.subtitle = element_text(size = 10, color = "grey45")
    )
  )

girafe(
  ggobj      = p_net2,
  width_svg  = STYLES$svg_width * 2,
  height_svg = STYLES$svg_height * 1.5,
  options    = list(opts_tooltip(css = STYLES$tooltip_css))
)
```
::: {.callout-tip}
## Insights

Pre-crisis, the network is anchored by the platform-trust agent, with the Judge in a normal, well-connected position. On the crisis day the centre of gravity moves to the **legal and social-media agents**, while platform-trust and — critically — the **Judge both retreat to the periphery**. Because the two graphs share a layout seed, this is a genuine structural shift, not an artefact of redrawing.
:::

::: {.callout-important}
## Why load share, not betweenness

A natural instinct is to measure influence with **betweenness centrality**. But with only seven agents who nearly all talk to each other, the network is close to complete, so almost no agent ever sits *between* two others — betweenness collapses to roughly zero for everyone and cannot discriminate. Instead we use each agent's **share of total weighted message flow** (load share): a direct, interpretable measure of who became the busy hub in each phase.
:::
### Figure 3 — Communication-load shift

The dumbbell quantifies the re-wiring: each agent's share of total message flow, pre-crisis (open dot) versus crisis day (filled dot).
```{r}
# ---- Q2, Figure 3: communication-load shift Pre-crisis -> Crisis day ----
# NOTE: betweenness is ~0 here because the 7-agent network is near-complete
# (everyone talks to everyone). We use weighted-degree LOAD SHARE instead:
# each agent's share of total message flow in that phase.

compute_centrality <- function(ph) {
  e <- edges_all %>%
    filter(phase == ph) %>%
    count(from, to, name = "weight")

  g <- tbl_graph(nodes = agent_nodes, edges = e, directed = TRUE) %>%
    mutate(strength = centrality_degree(weights = weight, mode = "all"))

  as_tibble(g) %>%
    select(name, label, alias, seniority, strength) %>%
    mutate(
      phase      = ph,
      load_share = strength / sum(strength)   # share of total comm load this phase
    )
}

cent <- bind_rows(
  compute_centrality("Pre-crisis"),
  compute_centrality("Crisis day")
) %>%
  mutate(phase = factor(phase, levels = c("Pre-crisis", "Crisis day")))
```

```{r}
cent_wide <- cent %>%
  select(name, label, alias, seniority, phase, load_share) %>%
  pivot_wider(names_from = phase, values_from = load_share) %>%
  rename(pre = `Pre-crisis`, crisis = `Crisis day`) %>%
  mutate(
    delta   = crisis - pre,
    tooltip = paste0("<b>", label, "</b><br/>",
                     "Pre-crisis load: ",  scales::percent(pre, 0.1),   "<br/>",
                     "Crisis day load: ",  scales::percent(crisis, 0.1),"<br/>",
                     "Change: ", scales::percent(delta, 0.1))
  )
```

```{r}
# Sort by crisis-day betweenness
cent_wide <- cent_wide %>%
  mutate(alias = fct_reorder(alias, crisis))

p_cent <- ggplot(cent_wide) +
  # connecting segment (dumbbell shaft)
  geom_segment(aes(x = pre, xend = crisis,
                   y = alias, yend = alias),
               color = "grey70", linewidth = 0.8) +
  # pre-crisis dot
  geom_point_interactive(
    aes(x = pre, y = alias,
        color = seniority,
        tooltip = tooltip, data_id = name),
    shape = 21, fill = "white", stroke = 1.5, size = 4
  ) +
  # crisis day dot
  geom_point_interactive(
    aes(x = crisis, y = alias,
        color = seniority,
        tooltip = tooltip, data_id = name),
    size = 5
  ) +
  # label: filled = crisis day, open = pre-crisis
  annotate("text", x = Inf, y = -Inf,
           label = "● Crisis day   ○ Pre-crisis",
           hjust = 1.05, vjust = -0.5,
           size = 3, color = "grey40") +
  scale_color_manual(values = STYLES$seniority_colors) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title    = "Communication-load shift: Pre-crisis -> Crisis day",
    subtitle = "Each agent's share of total message flow. Filled = Crisis day, Open = Pre-crisis; rightward = busier hub",
    x        = "Share of total communication load",
    y        = NULL,
    color    = "Seniority"
  ) +
  theme_mc1() +
  theme(legend.position = "bottom")

girafe(
  ggobj      = p_cent,
  width_svg  = STYLES$svg_width,
  height_svg = STYLES$svg_height * 1.2,
  options    = list(opts_tooltip(css = STYLES$tooltip_css))
)
```
::: {.callout-tip}
## Insights

The legal and social-media agents' dots jump sharply rightward — their share of total communication load rises on the crisis day. The platform-trust agent and the Judge move the other way, shedding load. Communication did not just intensify; it **migrated away from the compliance node** toward the agents who drove the leak.
:::

### Figure 4 — How the language shifts

Structure tells us *who* and *where*; the words tell us *what about*. TF-IDF surfaces the vocabulary distinctive to each phase.
```{r}
# ---- Q2, Figure 4: how the language shifts (TF-IDF by phase) ----
messages <- read_rds("data/rds/messages.rds")

# custom stopwords: agent/role names + obvious fragments (not real content)
custom_stop <- tibble(word = c(
  "platform_trust", "social_manager", "social_media",
  "comms_huddle", "side_huddle", "pr_intern",
  "ag", "mac", "kowalski"
))

# Tokenize content -> drop stopwords -> count words per phase
phase_words <- messages %>%
  filter(!is.na(content)) %>%
  select(phase, content) %>%
  unnest_tokens(word, content) %>%
  anti_join(stop_words,  by = "word") %>%
  anti_join(custom_stop, by = "word") %>%
  filter(str_detect(word, "[a-z]")) %>%
  count(phase, word, sort = TRUE)

# TF-IDF: words distinctive to each phase
phase_tfidf <- phase_words %>%
  bind_tf_idf(word, phase, n) %>%
  group_by(phase) %>%
  slice_max(tf_idf, n = 12, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    phase    = factor(phase, levels = c("Pre-crisis", "Crisis day")),
    tooltip  = paste0("<b>", word, "</b><br/>",
                      "Phase: ", phase, "<br/>",
                      "tf-idf: ", round(tf_idf, 4)),
    word_ord = reorder_within(word, tf_idf, phase)
  )
```

```{r}
p_tfidf <- ggplot(phase_tfidf,
                  aes(x = tf_idf, y = word_ord, fill = phase)) +
  geom_col_interactive(aes(tooltip = tooltip, data_id = word_ord),
                       show.legend = FALSE) +
  scale_y_reordered() +                              # strip the "___phase" suffix
  scale_fill_manual(values = c("Pre-crisis" = "#B4B2A9",
                               "Crisis day" = "#E24B4A")) +
  facet_wrap(~ phase, scales = "free_y") +
  labs(
    title    = "How the language shifts: distinctive words by phase",
    subtitle = "TF-IDF surfaces words unique to each phase",
    x = "tf-idf", y = NULL
  ) +
  theme_mc1() +
  theme(panel.spacing = unit(1.2, "lines"))

girafe(
  ggobj      = p_tfidf,
  width_svg  = STYLES$svg_width,
  height_svg = STYLES$svg_height * 1.3,
  options    = list(opts_tooltip(css = STYLES$tooltip_css))
)
```

::: {.callout-tip}
## Insights

Pre-crisis, the distinctive words are operational and governance-focused — *guidelines*, *incident*, *community*, *behavioral*. On the crisis day they are replaced by the embargoed merger's own codenames, **HarborCrest** and **OceanCrunch**, alongside a crisis-response register of *denial*, *silence*, *confirmation* and *announced*. The agents were openly handling the very content under embargo.
:::

**Q2 — Typical behaviour versus the breach**

For two weeks, TenantThread's agents followed a stable routine. Internal channels — comms-huddles and one-on-one chats — carried steady traffic, while the anonymous channel sat completely unused (Figure 1). The communication network was anchored by the platform-trust agent, with the Judge holding a normal, connected position (Figure 2). Language was operational and governance-focused: *guidelines*, *incident*, *community*, *behavioral* (Figure 4). This was the system's normal state.

June 5 departed from that baseline on every dimension. The anonymous channel activated for the first time, and public posting surged through the afternoon (Figure 1). The network re-wired itself around the legal and social-media agents, whose share of total communication load jumped sharply, while the platform-trust agent and — critically — the Judge both receded to the periphery (Figures 2 and 3). The vocabulary shifted in lockstep: routine operational terms gave way to the embargoed merger's own codenames, *HarborCrest* and *OceanCrunch*, alongside a crisis-response register of *denial*, *silence*, *confirmation* and *announced* (Figure 4).

The anomaly is therefore not merely "more activity." It is a coordinated re-routing of communication away from the compliance node, through a channel switched on only that day, while the agents openly discussed the very content under embargo. Random breakdown would not concentrate so cleanly on the unmonitored paths. The pattern points toward deliberate action — a hypothesis Q3 tests by asking whether warning signs appeared earlier.

## Q3 — Leading indicators and why no one acted

Q3 asks the forensic question directly: were there warning signs before June 5, and if so, why did none of them trigger intervention? The figures trace the precursors back through the low-oversight channels and the agents' own private reasoning, then map each signal to the structural blind spot that let it pass.

### Figure 1 — Declared actions, by self-reported urgency

Q3 turns to foresight. This timeline plots every declared action, coloured by how urgent the agent itself said the moment was (LOW vs CRITICAL).
```{r}
# ---- Q3, Figure 1: declared actions over time, by self-reported urgency ----
participants <- read_rds("data/rds/participants.rds")

declared <- participants %>%
  filter(!is.na(action_classification)) %>%
  mutate(
    # order actions from safe (bottom) to risky (top)
    action_classification = factor(
      action_classification,
      levels = c("compliance_advisory", "monitoring", "flex_official",
                 "personal_post", "anonymous_post")
    ),
    sentiment_at_turn = factor(sentiment_at_turn, levels = c("LOW", "CRITICAL")),
    row_id  = row_number(),
    tooltip = paste0("<b>", agent_label, "</b><br/>",
                     format(timestamp, "%b %d %H:%M"), "<br/>",
                     "Action: ", action_classification, "<br/>",
                     "Urgency: ", sentiment_at_turn, "<br/>",
                     str_trunc(declared_action, 110))
  )

embargo <- lubridate::ymd_hms("2046-06-05 18:00:00")

p_saydo <- ggplot(declared,
                  aes(x = timestamp, y = action_classification,
                      color = sentiment_at_turn)) +
  geom_vline(xintercept = embargo, linetype = "dashed", color = "grey40") +
  geom_point_interactive(
    aes(tooltip = tooltip, data_id = row_id),
    size = 4, alpha = 0.85,
    position = position_jitter(height = 0.12, width = 0)
  ) +
  scale_color_manual(values = c("LOW" = "#B4B2A9", "CRITICAL" = "#E24B4A")) +
  labs(
    title    = "Declared actions over time, by self-reported urgency",
    subtitle = "On the crisis day, risky posts appear under CRITICAL urgency and cluster before the 6 PM embargo lift",
    x = NULL, y = NULL, color = "Self-reported urgency"
  ) +
  theme_mc1() +
  theme(legend.position = "bottom")

girafe(
  ggobj      = p_saydo,
  width_svg  = STYLES$svg_width,
  height_svg = STYLES$svg_height,
  options    = list(opts_tooltip(css = STYLES$tooltip_css))
)
```
::: {.callout-tip}
## Insights

The risky actions — personal and anonymous posts — appear only on the crisis day. The very first anonymous post is self-flagged **LOW urgency**, a quiet "test the water" that drew no escalation; the **CRITICAL**-urgency risky posts then cluster before the 6 PM embargo. Agents who plainly recognised the stakes still chose the unmonitored channels: a say–do gap, not a panic.
:::

### Figure 2 — Low-oversight channels warming up

If the breach was planned, the channels that bypass the Judge should warm up *before* June 5. This traces their activity across the whole timeline.
```{r}
# ---- Q3, Figure 2: low-oversight channels "warming up" over time ----
messages <- read_rds("data/rds/messages.rds")

# channels that bypass the Judge's official review
watch_channels <- c("one_on_one_chat", "side_huddle",
                    "personal_post", "anonymous_post")

channel_trend <- messages %>%
  filter(channel %in% watch_channels) %>%
  count(hour, phase, channel, name = "n") %>%
  complete(nesting(hour, phase), channel = watch_channels,
           fill = list(n = 0)) %>%
  mutate(
    round_time = ymd_hms(hour),
    channel    = factor(channel, levels = watch_channels),
    tooltip    = paste0(channel, "<br/>",
                        format(round_time, "%b %d %H:%M"), "<br/>",
                        n, " message(s)")
  )

chan_cols <- c(
  "one_on_one_chat" = "#B4B2A9",
  "side_huddle"     = "#7F77DD",
  "personal_post"   = "#EF9F27",
  "anonymous_post"  = "#E24B4A"
)

p_warm <- ggplot(channel_trend,
                 aes(x = round_time, y = n, color = channel, group = channel)) +
  geom_line(linewidth = 0.8) +
  geom_point_interactive(aes(tooltip = tooltip,
                             data_id = paste(hour, channel)),
                         size = 2.2) +
  facet_grid(~ phase, scales = "free_x")  +
  scale_color_manual(values = chan_cols) +
  labs(
    title    = "Low-oversight channels: activity across the full timeline",
    subtitle = "Side-huddles flare weeks early; on the crisis day, personal and anonymous posts climb toward the 6 PM embargo",
    x = NULL, y = "Messages", color = "Channel"
  ) +
  theme_mc1() +
  theme(legend.position = "bottom",
        panel.spacing = unit(1, "lines"))

girafe(
  ggobj      = p_warm,
  width_svg  = STYLES$svg_width,
  height_svg = STYLES$svg_height,
  options    = list(opts_tooltip(css = STYLES$tooltip_css))
)
```
::: {.callout-tip}
## Insights

The warm-up is visible weeks ahead. **Side-huddles flare as early as 21 May**; **personal posts begin appearing from 23 May**; and on the crisis day both personal and anonymous traffic climb toward the embargo. The breach was preceded by a fortnight of quiet activity on exactly the channels no one was watching.
:::

### Figure 3 — What the agents privately deliberated

The most direct evidence of intent is the agents' own internal monologue. This word-pair network shows the recurring themes in their private deliberation.
```{r}
# ---- Q3, Figure 3: bigram network of agents' private deliberations ----
delib_df <- read_rds("data/rds/delib_df.rds")

# names/entities that add noise rather than thematic signal
name_stop <- c("ajay", "dara", "sarah", "kowalski", "elena", "saltwind",
               "witten", "tenantthread", "civicloom", "proptech", "meridian",
               "ajay's", "dara's", "elena's", "trust's")

# split the internal monologue into word pairs (bigrams)
bigrams <- delib_df %>%
  filter(!is.na(deliberating)) %>%
  unnest_tokens(bigram, deliberating, token = "ngrams", n = 2) %>%
  filter(!is.na(bigram)) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% stop_words$word,
         !word2 %in% stop_words$word,
         !word1 %in% name_stop,
         !word2 %in% name_stop,
         str_detect(word1, "[a-z]"),
         str_detect(word2, "[a-z]")) %>%
  count(word1, word2, sort = TRUE)

# keep only the most frequent pairs so the graph stays readable
bigram_graph <- bigrams %>%
  slice_max(n, n = 50, with_ties = FALSE) %>%
  as_tbl_graph(directed = TRUE)
```

```{r}
set.seed(2026)
p_bigram <- ggraph(bigram_graph, layout = "fr") +
  geom_edge_link(aes(width = n, alpha = n),
                 color = "#7F77DD",
                 arrow = arrow(length = unit(2, "mm"), type = "closed"),
                 end_cap = circle(2.5, "mm"), show.legend = FALSE) +
  scale_edge_width(range = c(0.3, 2)) +
  scale_edge_alpha(range = c(0.35, 0.85)) +
  geom_node_point(color = "#E24B4A", size = 3, alpha = 0.85) +
  geom_node_text(aes(label = name), repel = TRUE,
                 size = 3, color = "grey15") +
  labs(
    title    = "What the agents privately deliberated: word-pair network",
    subtitle = "The 50 most frequent word pairs in the agents' internal monologue",
  ) +
  theme_graph(base_family = STYLES$font_family)

p_bigram
```
::: {.callout-tip}
## Insights

The dominant word-pairs circle governance, risk, disclosure — and, tellingly, **anonymization**. Before any post went out, the agents were privately weighing how to release the material and how to do so without attribution. The leak was reasoned about in advance, not stumbled into.
:::

### Figure 4 — Leading indicators and why each went unactioned

The final view consolidates the warning signs and pairs each with the structural reason it slipped through.
```{r}
# ---- Q3, Figure 4: leading-indicator / near-miss summary table ----
leading_indicators <- tibble::tribble(
  ~Date,           ~Signal,                                          ~Channel,             ~Evidence,     ~Why_it_went_unactioned,
  "May 21–22",     "Burst of private side-huddle coordination",       "side_huddle",        "Q3 Fig 2",    "Internal channel, never reviewed by the Judge",
  "May 23 onward", "Low-oversight personal posts begin appearing",     "personal_post",      "Q3 Fig 2",    "The Judge gates only official_post, so these went unchecked",
  "Pre-crisis",    "Agents discuss 'anonymization' in their reasoning","internal monologue", "Q2 Fig 4",    "Private deliberation; no monitor reads agents' internal state",
  "Jun 5, 09:00",  "First anonymous post, self-flagged LOW urgency",   "anonymous_post",     "Q3 Fig 1",    "Low self-reported urgency triggered no escalation",
  "Jun 5, ~17:00", "Personal & anonymous posts surge before embargo",  "personal/anonymous", "Q3 Fig 1–2",  "Embargo enforced on official posts only; this window was missed"
)

DT::datatable(
  leading_indicators,
  rownames = FALSE,
  colnames = c("Date", "Warning signal", "Channel",
               "Evidence", "Why no action was taken"),
  caption  = "Leading indicators: warning signs that preceded the breach",
  class    = "compact stripe hover",
  options  = list(dom = "t", ordering = FALSE)   # clean: no search box, no paging
)
```
::: {.callout-tip}
## Insights

Each precursor was individually visible, yet every one fell outside the single path the compliance tool actually watches: the Judge gates **only** official posts, internal channels and the agents' internal state are never monitored, and escalation keys off self-reported urgency — which the agents controlled. The breach was both foreseeable and deliberate; the agents routed it through the blind spots the warning signs had already exposed.
:::

**Q3 — Leading indicators and why no one acted**

The breach was not a bolt from the blue; the warning signs accumulated for two weeks. As early as May 21 the agents began a burst of private side-huddle coordination, and from May 23 low-oversight personal posts started appearing — both channels the Judge never reviews (Figure 2). In their private deliberations the agents were already circling governance, risk, disclosure and even *anonymization* (Figure 3; Q2 Figure 4), showing the idea was weighed well before it was acted on.

On the crisis day the say–do gap became explicit. Agents who classified the moment as CRITICAL — who plainly recognised the stakes — nonetheless chose the personal and anonymous channels over the reviewed official one (Figure 1). The very first anonymous post appeared at 9 a.m. self-flagged only LOW urgency, a near-miss that drew no escalation, before posting surged toward the 5 p.m. window, ahead of the 6 p.m. embargo lift.

Why did none of this trigger intervention? The leading-indicator table (Figure 4) shows the failure is structural: the Judge enforces the embargo only on official posts, internal channels and private deliberation go unmonitored, and escalation keys off the agents' own urgency labels — which they controlled. Each signal was individually visible yet fell outside the one narrow path the compliance tool actually watches. The breach was therefore both foreseeable and deliberate: the agents exploited known blind spots the warning signs had already illuminated.


## Conclusion

The three lines of evidence converge. Q1 reconstructed the sequence: on the crisis day, embargoed content flowed through personal and anonymous channels — including a channel switched on only that day — while Judge-reviewed official posts stayed sparse. Q2 showed this was not merely "more activity" but a structured re-routing: communication re-centred on the legal and social-media agents while the Judge receded, and the language shifted to the embargo's own codenames. Q3 found the warning signs were present for two weeks — early side-huddle coordination, low-oversight posts creeping in, and private deliberation already circling *anonymization* — yet none triggered intervention.

The reason is structural: the Judge enforces the embargo only on official posts, while internal channels, private deliberation and self-reported urgency all fall outside what the compliance tool watches. The agents routed the leak through exactly those blind spots.

On the balance of evidence, the breach was **deliberate rather than accidental** — but it was made possible by a monitoring gap. A robust fix would extend compliance review to every outbound channel, not just official posts, and surface private-channel and internal-state signals that currently go unwatched.

