---
title: "Profitability & Volatility Drivers"
subtitle: "A Visual Analytics Study of a Motor Insurance Portfolio"
author: "Wang Xiaohan"
date: "May 21, 2026"
format:
  revealjs:
    theme: [simple, slides-custom.scss]
    slide-number: true
    transition: slide
    transition-speed: slow
    background-transition: fade
    incremental: false
    code-fold: true
execute:
  echo: false
  warning: false
  freeze: auto
---
```{r}
#| label: setup
#| include: false
pacman::p_load(tidyverse, lubridate, ggdist, ggridges, ggthemes,
               colorspace, ggstatsplot, patchwork, skimr, broom, scales, ggcorrplot)

raw <- read_delim("data/Motor vehicle insurance data.csv", delim = ";",
                  show_col_types = FALSE)

dat <- raw %>%
  mutate(across(c(Date_start_contract, Date_last_renewal, Date_next_renewal,
                  Date_birth, Date_driving_licence, Date_lapse), ~ dmy(.x))) %>%
  mutate(
    Type_risk = factor(Type_risk, levels = c(1,2,3,4),
      labels = c("Motorbike","Van","Passenger car","Agricultural")),
    Area = factor(Area, levels = c(0,1), labels = c("Rural","Urban")))

sandbox <- dat %>%
  mutate(
    Loss_Ratio = if_else(Premium > 0, Cost_claims_year / Premium, NA_real_),
    Has_Claim  = factor(if_else(N_claims_year > 0, "Claim", "No claim")))

lr_cap <- quantile(sandbox$Loss_Ratio, 0.99, na.rm = TRUE)
sandbox <- sandbox %>% mutate(Loss_Ratio_capped = pmin(Loss_Ratio, lr_cap))
```

## Contents {.smaller .contents-slide}

1. The business question & dataset
2. How losses are distributed
3. Do vehicle types differ?
4. Are categorical factors material?
5. Correlations among risk factors
6. Fair comparison: the funnel plot
7. What really drives a claim?
8. Key findings & implications

::: {.notes}
This executive summary investigates what drives the profitability and volatility of a Spanish motor insurance portfolio. The analysis moves from describing how losses are distributed, through statistically testing candidate factors, to a multivariable model and a fair-comparison funnel plot, ending with concrete underwriting implications.
:::

## The Business Question {.smaller}

:::: {.columns}
::: {.column width="52%"}
> *"Insurers often carry underperforming portfolios where the root causes of poor profitability and high volatility are unknown."*

This study interrogates a real-world Spanish motor insurance portfolio to find those root causes — and to **prove them statistically**, not by intuition.
:::
::: {.column width="48%"}
::: {.bignum}
105,555
:::
::: {.bignum-label}
policy-year records · 30 variables · Spain, 2015–2018
:::

<br>

- No ready-made profitability field → I derive the **Loss Ratio** (claims ÷ premium)
- Above **100%** = a technical loss
- **Goal:** validate drivers of **profitability** & **volatility**
:::
::::

::: {.notes}
The portfolio contains over 100,000 policy-year records across 30 variables. Because the dataset has no ready-made profitability field, the central metric — the Loss Ratio, claims divided by premium — is derived during data preparation. A value above 100% indicates a technical loss. All subsequent analysis revolves around explaining this measure.
:::

## How Losses Are Distributed {.smaller}

:::: {.columns}

::: {.column width="68%"}
```{r}
#| fig-width: 8.5
#| fig-height: 5.2
sandbox %>%
  filter(Has_Claim == "Claim", !is.na(Loss_Ratio)) %>%
  mutate(LR_plot = pmin(Loss_Ratio, 3)) %>%      
  ggplot(aes(x = LR_plot)) +
  geom_histogram(bins = 50, fill = "#e36c52", colour = "white", alpha = 0.95) +
  geom_vline(xintercept = 1, colour = "#1a1a1a",
             linewidth = 0.9, linetype = "dashed") +
  annotate("text", x = 1.02, y = Inf, label = "Break-even (100%)",
           hjust = 0, vjust = 1.6, size = 4.2, colour = "#1a1a1a") +
  annotate("text", x = 3, y = Inf, label = "→ capped at 300%\n(heavy tail beyond)",
           hjust = 1, vjust = 1.4, size = 3.6, colour = "#9c3d27") +
  scale_x_continuous(labels = scales::percent_format(),
                     breaks = seq(0, 3, 0.5)) +
  labs(x = "Loss Ratio (Claims / Premium)",
       y = "Number of claiming policies") +
  theme_minimal(base_size = 15) +
  theme(panel.grid.minor = element_blank(),
        axis.title = element_text(face = "bold"))
```
:::

::: {.column width="32%"}
::: {.dist-read}
**How to read it**

Each bar = policies at that loss level. The dashed line is **break-even** — left of it, the policy made money; right of it, claims beat premium.

The bulk sits **left of break-even**, but a long **right tail** keeps going well past 300%.
:::
:::

::::

::: {.dist-points}
**81%** of policies never claim · among claimers, **most stay below break-even** · a **heavy tail** runs at a technical loss → **profitability lives in the tail, not the average**
:::

::: {.notes}
About 81% of policies have no claim at all. Among those that do claim, the Loss Ratio is strongly right-skewed: the bulk of claiming policies sit below the 100% break-even line, meaning their claims cost less than the premium collected. But a long right tail extends far beyond break-even — I've capped the axis at 300% here so the main distribution stays readable, but in the raw data the tail runs much further, up to catastrophic policies whose losses exceed eight times the premium. The key insight for underwriting is that aggregate profitability is driven by this small, extreme tail rather than by the typical policy. This argues for tail-focused risk segmentation rather than across-the-board rate increases.
:::
::::

## Do Vehicle Types Differ? {.smaller}

:::: {.columns}
::: {.column width="65%"}
```{r}
#| fig-width: 6.5
#| fig-height: 4.5
sandbox %>%
  filter(Has_Claim == "Claim", !is.na(Loss_Ratio_capped)) %>%
  ggplot(aes(x = Loss_Ratio_capped, y = Type_risk, fill = factor(stat(quantile)))) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE,
                      quantiles = 4, quantile_lines = TRUE) +
  scale_fill_viridis_d(name = "Quartile") +
  scale_x_continuous(labels = scales::percent_format(),
                     breaks = seq(0, 8, 2)) +
  labs(x = "Loss Ratio (Claims / Premium, capped)", y = NULL) +
  theme_ridges(font_size = 13) +
  theme(panel.grid.major = element_line(colour = "grey90"),
        axis.title.x = element_text(face = "bold", hjust = 0.5))
```
:::
::: {.column width="35%"}
- All four vehicle types share a similar peak near **Loss Ratio = 0**
- Motorbikes & vans show **heavier right tails** than agricultural
- Quartile colours show wide spread (**Q4 reaches the cap**)
- Shapes **look** distinct — but is the difference **statistically real**?
:::

::::

::: {.notes}
The ridgeline plot compares Loss Ratio distributions across the four vehicle types, with fill encoding quartiles. Motorbikes and vans show heavier right tails than agricultural vehicles. These shapes look visually distinct — but as the next slide shows, that visual difference does not survive a formal statistical test.
:::


## A Formal Test Says: No {.smaller}

:::: {.columns}
::: {.column width="55%"}
```{r}
#| fig-width: 6
#| fig-height: 5
set.seed(1234)
br_dat <- sandbox %>%
  filter(Has_Claim == "Claim", !is.na(Loss_Ratio), Loss_Ratio < 5) %>%
  slice_sample(n = 3000)
ggbetweenstats(data = br_dat, x = Type_risk, y = Loss_Ratio,
               type = "nonparametric", xlab = "Vehicle type", ylab = "Loss Ratio",
               results.subtitle = TRUE)
```
:::
::: {.column width="45%"}
::: {.bignum}
p = 0.32
:::
::: {.bignum-label}
Kruskal-Wallis · not significant
:::

<br>

The ridgelines *looked* different — but the medians barely move: **0.60 · 0.64 · 0.76**.

Vehicle type does **not** drive claim **severity**.

> So the structure must lie in **frequency** — *who* files a claim, not *how much*.
:::
::::

::: {.notes}
A Kruskal-Wallis test returns p = 0.32, well above 0.05. Despite the visually distinct ridgeline shapes, vehicle type does not significantly affect claim severity once a claim has occurred. This negative result is important: it redirects the search for drivers away from severity towards claim frequency. Honest reporting of a non-significant result reflects sound statistical practice.
:::

## The Large-Sample Trap {.smaller}

```{r}
#| fig-width: 11
#| fig-height: 3.5
ggbarstats(data = sandbox, x = Has_Claim, y = Area,
           xlab = "Area", legend.title = "Claim status",
           results.subtitle = TRUE)
```
:::: {.trap-cards}

::: {.trap-card}
**Statistically significant**

χ² = **86.4**, p = **1.5e-20**

A textbook "highly significant" result.
:::

::: {.trap-card}
**Practically negligible**

Cramer's V = **0.03** · gap is only **18% vs 20%**

The association is essentially zero.
:::

::: {.trap-card .trap-card-accent}
**Why? Sample size.**

With **105,555** rows, even trivial gaps cross the significance line.

→ **Effect size matters, not just the p-value.**
:::

::::

::: {.notes}
The chi-square test is highly significant, suggesting area matters. But the effect size, Cramer's V, is only 0.03 — negligible. The actual gap is just 18% rural versus 20% urban. With over 100,000 records, even a trivial difference becomes statistically significant. Reading the p-value alone would mislead; the effect size shows area is not a material driver.
:::


## No Strong Numeric Predictor {.smaller}

:::: {.columns}
::: {.column width="58%"}
```{r}
#| fig-width: 5.5
#| fig-height: 4.5
num_vars <- sandbox %>%
  select(Premium, Cost_claims_year, Power, Cylinder_capacity,
         Value_vehicle, Weight) %>%
  drop_na()
ggcorrplot(cor(num_vars, method = "spearman"), type = "lower", lab = TRUE,
           lab_size = 3, colors = c("#2166ac", "white", "#b2182b"))
```
:::
::: {.column width="42%"}
- Vehicle attributes are **highly intercorrelated**: Value–Weight **0.87**, Cylinder–Value **0.79**
- Claim cost correlates **weakly with all factors** (|ρ| ≤ **0.09**)
- Premium tracks vehicle attributes (**0.39–0.44**), not actual loss
- **No single variable predicts loss** → risk is idiosyncratic
:::
::::

::: {.notes}
The heatmap shows two things. Vehicle attributes are highly intercorrelated — value, weight, power and cylinder capacity all measure the same underlying size. More importantly, claim cost correlates only weakly with every factor. No single numeric variable predicts loss; risk in this portfolio is largely idiosyncratic, which is why a multivariable approach is needed.
:::

## A Fair Comparison: The Funnel Plot {.smaller}

::::: {.columns}

:::: {.column width="60%"}
```{r}
#| fig-width: 6
#| fig-height: 4.5
pacman::p_load(FunnelPlotR)
seg <- sandbox %>%
  group_by(Type_risk, Area) %>%
  summarise(claims = sum(N_claims_year > 0), n = n(), .groups = "drop") %>%
  mutate(segment = paste(Type_risk, Area, sep = " - "))
funnel_plot(.data = seg, numerator = claims, denominator = n,
            group = segment, data_type = "PR",
            title = "Claim incidence by segment",
            x_label = "Exposure (number of policies)", y_label = "Claim rate")
```
::::

:::: {.column width="40%"}
::: {.funnel-block}
**Inside the funnel → noise**

Small segments (agricultural, **~850** policies) swing by chance and sit **deep inside** the limits. The largest segment (cars, **~60,000**) lands on the mean line — **stable**.
:::

::: {.funnel-block .funnel-block-accent}
**Outside the limits → real**

**2 of 8** segments fall outside the control limits — genuinely unusual, not sampling noise. Limits are **adjusted for overdispersion**, so the test is fair across segment sizes.
:::
::::

:::::

::: {.notes}
The funnel plot reframes the comparison correctly. The two agricultural segments sit far left with very low exposure — a naive ranking would call them the safest, but their position inside the wide funnel shows this is noise, not signal. Two segments fall outside the control limits and are genuinely high-risk. The largest segment lies on the mean line, confirming stability. The plot adjusts for overdispersion, appropriate for insurance claims.
:::



## What Really Drives a Claim? {.smaller}

:::: {.columns}
::: {.column width="60%"}
```{r}
#| fig-width: 6
#| fig-height: 4.5
mod_dat <- sandbox %>%
  mutate(Vehicle_Age = 2018 - Year_matriculation,
         Age = 2019 - year(Date_birth),
         Second_driver = factor(Second_driver, levels = c(0,1),
                                labels = c("No","Yes"))) %>%
  select(Has_Claim, Type_risk, Area, Value_vehicle, Power,
         Age, Vehicle_Age, Second_driver) %>%
  drop_na()
mod <- glm(Has_Claim ~ ., data = mod_dat, family = binomial)
tidy(mod, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term != "(Intercept)") %>%
  ggplot(aes(x = estimate, y = reorder(term, estimate))) +
  geom_point(colour = "#b2182b", size = 3) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.25) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_x_log10() +
  labs(x = "Odds Ratio (log scale)", y = NULL) +
  theme_minimal(base_size = 12)
```
:::
::: {.column width="40%"}
- **Vehicle type** is the one robust driver of claim incidence
- Vans & cars have **OR < 1** → lower odds than motorbikes
- Numeric factors (age, value, power) cluster at **OR ≈ 1** → little effect
- Agricultural **OR ≈ 100** but **CI > 1,000** → small-sample noise
- **Funnel plot & model cross-validate** each other
:::
::::

::: {.notes}
The forest plot delivers the clearest answer. After controlling for all factors, vehicle type is the one robust driver: vans and passenger cars have significantly lower claim odds than motorbikes. Numeric factors cluster around an odds ratio of one. Agricultural vehicles show a huge odds ratio but an extremely wide confidence interval — an unreliable small-sample estimate, exactly the artefact the funnel plot warned against. The two methods cross-validate each other.
:::

## Key Findings and Implications {.smaller}

::::: {.columns}

:::: {.column width="50%"}
::: {.kf-label}
What we found
:::

::: {.kf-finding}
**Profitability is concentrated** — about **91%** of policies are individually profitable, yet a small catastrophic tail (Technical Result down to **−€260,000**) dominates the portfolio's economics. Managing the tail matters far more than average pricing.
:::

::: {.kf-finding}
**No single factor predicts loss** — vehicle type (Kruskal-Wallis **p = 0.32**), area (Cramer's **V = 0.03**) and vehicle value (Spearman **ρ = 0.13**) are all statistically weak or negligible. Loss is largely **idiosyncratic**, not explained by the rating factors on file.
:::

::: {.kf-finding}
**Structure lies in frequency, not severity** — after controlling for all factors, vehicle type is the one robust driver of *whether* a claim occurs (vans & cars **OR < 1**). The dramatic "agricultural risk" (**OR ≈ 100**) was small-sample noise — confirmed by the funnel plot.
:::
::::

:::: {.column width="50%"}
::: {.kf-label}
What to do
:::

::: {.kf-action}
**1 · Manage the tail, not the average**

Focus underwriting and reinsurance on the rare catastrophic policies that drive losses, rather than raising rates across the whole book.
:::

::: {.kf-action}
**2 · Compare segments by exposure**

Judge segment risk on a funnel plot that widens for small samples — not on raw claim rates, which made tiny agricultural cohorts look falsely extreme.
:::

::: {.kf-action .kf-action-accent}
**3 · Enrich the data**

Existing rating factors miss true risk. Invest in richer features — driver behaviour, telematics, history — to capture the idiosyncratic variation current fields cannot.
:::
::::

:::::

## {.center .takeaway-slide}

::: {.takeaway-text}
The portfolio's fate is written in its **tail** and its **claim frequency** —
not in any single rating factor.
:::

::: {.takeaway-sub}
Manage the tail · compare by exposure · enrich the data
:::

::: {.notes}
The single most important takeaway: the portfolio's profitability and volatility are driven by the catastrophic tail and by claim frequency, not by any individual rating factor examined. The three actions follow directly — manage tail policies, compare segments fairly by exposure, and enrich the data to capture true risk.
:::